\LARGE \bf Planning with Attitude

### 2020-10-14 paper/main.tex
more small tweaks

    The Jacobian of $r(q)$ can be found using \eqref{eq:quat_gradient}:
    \begin{align} \label{eq:newton_foc}
            \nabla r(q) &= \pdv{r}{q} G(q) \\
              &= \left( \sum_i -2 H^T R(q)^T R({}^B \hat{v}_i) G(q)\right).
    \end{align}
    Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update, $\phi_k$ via the Moore-Penrose psuedoinverse:

@@

        A 180 degree barrel roll trajectory for a fixed-wing airplane was optimized. The
        airplane's dynamics model consists of the a simple rigid body as defined in
        Section \ref{sec:rigidbody_dynamics} with forces and torques due to lift and drag
        fit from wind tunnel data \cite{manchester2016udp}. The airplane was tasked to do
        a barrel roll by constraining the terminal state to upside-down (see Fig.
        \ref{fig:barrellroll}). To avoid issues with the double-cover and integration error,
        the following was used to enforce a terminal orientation to $\bar{q}$:
        \begin{equation}
            \frac{q_v}{\norm{q}} - \bar{q}_v
                \sign\left(\bar{q}^T \frac{q }{ \norm{q}}\right) = 0
        \end{equation}

@@

\section{Conclusions} \label{sec:conclusion}
    We have presented a general, unified method for optimization-based planning and control for rigid-body
    systems with arbitrary attitude using standard linear algebra and vector calculus.
    We have demonstrated that the application of this methodology is straightforward and
    yields substantial improvements in the convergence of Newton-based methods (see Fig. 
    \ref{fig:wahba_convergence}), while also offering improvements for nonlinear constrained
    trajectory optimization for floating-base systems (see Table \ref{tab:timing_results}).

    \todo{This is a really nice pitch. Can we say something like this in the intro?}
    With the modifications presented, ALTRO can solve problems few other methods for
    trajectory optimization can. Many state-of-the-art methods, such as direct
    collocation or sequential convex programming, rely on commercial, proprietary, or
    general-purpose solvers whose internal numerics often are abstracted away from the
    user. By exploiting the unique structure of both the trajectory optimization problem
    and the rotation group, ALTRO will likely be able to solve more challenging problems
    with performance.

    The methods presented here can easily be leveraged to adapt other classes
    of gradient or Newton-based algorithms to exploit the structure of 3D rotations. Future
    work may include adaptation of methods for state estimation, localization, design, or 
    other methods for motion planning such as direct collocation.

    Future work will focus on continued 
    refinement of ALTRO and performance improvements for use in real-time model-predictive control applications, as well as extensions to multi-body robotic
    systems, such as humanoids or quadrupeds, represented in
    ``maximal'' coordinates that include the 3D orientation of each body \cite{brudigam2020linear}.

\section*{Acknowledgements}
This work was supported by a NASA Early Career Faculty Award (Grant Number 80NSSC18K1503). This material is based upon work supported by the National Science Foundation Graduate
Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and
conclusions or recommendations expressed in this material are those of the author(s) and
do not necessarily reflect the views of the National Science Foundation.

>>>>

    The Jacobian of $r(q)$ can be found using \eqref{eq:quat_gradient}:
    \begin{align} \label{eq:newton_foc}
            \nabla r(q) &= \pdv{r}{q} G(q) \\
              &= -2 H^T R(q)^T \left( \sum_i R({}^B \hat{v}_i) \right) G(q) .
    \end{align}
    Given a guess solution, $q_k$, The standard Gauss-Newton method can then be used to compute a three-parameter update, $\phi_k$ via the Moore-Penrose psuedoinverse:

@@

        A 180 degree barrel roll trajectory for a fixed-wing airplane was optimized. The
        airplane's dynamics model consists of the a simple rigid body as defined in
        Section \ref{sec:rigidbody_dynamics} with forces and torques due to lift and drag
        fit from wind tunnel data \cite{manchester2016udp}. The airplane was tasked to do
        a barrel roll by constraining the terminal state to upside-down (see Fig.
        \ref{fig:barrellroll}). To mitigate issues with integration error and drift in the magnitude of the quaternion,
        the following constraint function was used to enforce a terminal orientation of $\bar{q}$:
        \begin{equation}
            \frac{q_v}{\norm{q}} - \bar{q}_v
                \sign\left(\bar{q}^T \frac{q }{ \norm{q}}\right) = 0
        \end{equation}

@@

\section{Conclusions} \label{sec:conclusion}
    We have presented a general, unified method for optimization-based planning and control for rigid-body
    systems with arbitrary attitude using standard linear algebra and vector calculus.
    The application of this methodology is straightforward and
    yields substantial improvements in the convergence of Newton and DDP-based methods, while also offering improvements for nonlinear constrained
    trajectory optimization for floating-base systems.

    Many state-of-the-art trajectory optimization methods, including direct collocation and sequential convex programming, rely on general-purpose optimization solvers whose internal numerical methods are not exposed to the user. Therefore, these methods are unable to exploit the full structure of both the trajectory optimization problem and the rotation group at a low level. In contrast, we are able to implement deep, native support for quaternions into the ALTRO solver, making it possible to solve more challenging problems with higher performance than other algorithms.

    In future work, we plan to apply ALTRO to real-time model-predictive control problems for aerial vehicles like quadrotors and airships that experience large attitude changes. The methods we have presented can also be leveraged to adapt other classes of gradient or Newton-based algorithms to exploit the structure of 3D rotations. Future directions beyond trajectory optimization may include simulation and planning methods that leverage maximal-coordinate formulations of multi-body dynamics~\cite{brudigam2020linear}, system-identification for complex multi-body systems and fixed-wing aircraft in post-stall conditions, and state estimation for spacecraft with sparse measurements.

\section*{Acknowledgements}
This work was supported by a NASA Early Career Faculty Award (Grant Number 80NSSC18K1503).
This research was carried out in part at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration and funded through JPL’s Strategic University Research Partnerships (SURP) program.
This material is based upon work supported by the National Science Foundation Graduate
Research Fellowship Program under Grant No. DGE-1656518. Any opinions, findings, and
conclusions or recommendations expressed in this material are those of the author(s) and
do not necessarily reflect the views of the National Science Foundation.
